---
title: "ADA Final Project"
author: "Leah Nason"
date: "11/27/2022"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Import Data 

View final project data and code at https://github.com/leahnason/ADA-Final-Project.

My syntax does not include raw GitHub links for RUCC and MICA datasets, which are .xls and .xlsx respectively.

```{r dataimport, warning=FALSE, message=FALSE}

# Load relevant packages
pacman::p_load(knitr, dplyr, tidyverse, stringr, writexl, readxl, kableExtra, summarytools, psych, utils, table1, MASS, stargazer, lmtest, broom, sandwich, stats)

# Set working directory
setwd("/Users/leahnason/Dropbox/My Mac (Leah’s MacBook Air)/Desktop/Fall 2022/Advanced Data Analysis/Final Project")

# Import data
MICA <- read_xlsx("Data/MICA_20221127.xlsx", col_names = TRUE, na = "x", skip = 2)
ACSSocial <- read_csv("https://raw.githubusercontent.com/leahnason/ADA-Final-Project/main/ACSDP5Y2018.DP02-2022-11-27T212858_social.csv", col_names = FALSE)
ACSEcon <- read_csv("https://raw.githubusercontent.com/leahnason/ADA-Final-Project/main/ACSDP5Y2018.DP03-2022-11-27T213313_econ.csv", col_names = FALSE)
ACSDemo <-read_csv("https://raw.githubusercontent.com/leahnason/ADA-Final-Project/main/ACSDP5Y2018.DP05-2022-11-27T213004_demo.csv", col_names = FALSE)
RUCC <- read_xls ("Data/ruralurbancodes2013.xls")

MICA <- MICA[-c(1,2), ] # remove first two blank rows
MICA <- MICA[-c(116:127), ] # remove data for "Total for Selection," "Missouri," and table comments

```

# 2. Clean MICA Data

```{r datacleanMICA, warning=FALSE, message=FALSE}

# MICA data includes the count and incidence of late-stage female breast cancer in Missouri

MICA <- MICA %>%
  dplyr::select(County, TotalCount, TotalRate) %>%
  drop_na()

MICA$TotalRate <- round(MICA$TotalRate) # create integer for age-adjusted Poisson regression

# Correct St. Louis City and St. Louis County observation for full merge
MICA$County <- sub("St. Louis County", "St. Louis", MICA$County) # remove "county" for merge
MICA$County <- sub("St. Louis City", "St. Louis city", MICA$County) # adjust capitalization for merge

```

# 3. Clean ACS Social Characteristics Data

```{r datacleanACSSocial, warning=FALSE, message=FALSE}

# ACS Social data includes education and household internet access 

ACSSocial <- as.data.frame(t(ACSSocial)) # transpose data frame

# Select percent 
ACSSocial <- ACSSocial %>%
  dplyr::select(County = V1, HighSchoolPlusPercent = V75, HouseholdInternetPercent = V171) %>%
  filter(!str_detect(County, "!!Estimate")) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error"))

ACSSocial$HouseholdInternetPercent <- sub("%", "", ACSSocial$HouseholdInternetPercent) 
# remove "%" in column for merge

ACSSocial$HighSchoolPlusPercent <- sub("%", "", ACSSocial$HighSchoolPlusPercent) 
# remove "%" in column for merge

ACSSocial <- ACSSocial[-c(1), ] # remove first row with variable description

ACSSocial$County <- sub(" County, Missouri!!Percent", "", ACSSocial$County) # prep for merge
ACSSocial$County <- sub(", Missouri!!Percent", "", ACSSocial$County) # prep for merge

```

# 4. Clean ACS Economic Characteristics Data

```{r datacleanACSEcon, warning=FALSE, message=FALSE}

# ACS Economic data includes household income and health insurance coverage 

ACSEcon <- as.data.frame(t(ACSEcon)) # transpose data frame

# Select estimate
ACSEconIncome <- ACSEcon %>%
  dplyr::select(County = V1, HouseholdMedianIncome = V69) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent"))

ACSEconIncome$HouseholdMedianIncome <- sub(",", "", ACSEconIncome$HouseholdMedianIncome) 
# remove "," in column

ACSEconIncome <- ACSEconIncome[-c(1), ] # remove first row with variable description

ACSEconIncome$County <- sub(" County, Missouri!!Estimate", "", ACSEconIncome$County) # prep for merge
ACSEconIncome$County <- sub(", Missouri!!Estimate", "", ACSEconIncome$County) # prep for merge

# Select percent
ACSEconInsurance <- ACSEcon %>%
  dplyr::select(County = V1, HealthInsurancePercent = V118) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error")) %>%
  filter(!str_detect(County, "!!Estimate"))

ACSEconInsurance$HealthInsurancePercent <- sub("%", "", ACSEconInsurance$HealthInsurancePercent) 
# remove "%" in column 

ACSEconInsurance <- ACSEconInsurance[-c(1), ] # remove first row with variable description

ACSEconInsurance$County <- sub(" County, Missouri!!Percent", "", ACSEconInsurance$County) # prep for merge
ACSEconInsurance$County  <- sub(", Missouri!!Percent", "", ACSEconInsurance$County) # prep for merge

ACSEcon <- merge(ACSEconIncome, ACSEconInsurance, by.x = "County", by.y = "County")

```

# 5. Clean ACS Demographics Data

```{r datacleanACSDemo, warning=FALSE, message=FALSE}

# ACS Demogrphics data includes gender, age, race/ethnicity

ACSDemo <- as.data.frame(t(ACSDemo)) # transpose data frame

# Select estimate
ACSDemoAge <- ACSDemo %>%
  dplyr::select(County = V1, MedianAge = V20) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent"))

ACSDemoAge  <- ACSDemoAge[-c(1), ] # remove first row with variable description

ACSDemoAge$County <- sub(" County, Missouri!!Estimate", "", ACSDemoAge$County) # prep for merge
ACSDemoAge$County <- sub(", Missouri!!Estimate", "", ACSDemoAge$County) # prep for merge

# Select percent
ACSDemoGender <- ACSDemo %>%
  dplyr::select(County = V1, FemalePercent = V33) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error")) %>%
  filter(!str_detect(County, "!!Estimate"))

ACSDemoGender$FemalePercent <- sub("%", "", ACSDemoGender$FemalePercent) 
# remove "%" in column 

ACSDemoGender <- ACSDemoGender[-c(1), ] # remove first row with variable description

ACSDemoGender$County <- sub(" County, Missouri!!Percent", "", ACSDemoGender$County) # prep for merge
ACSDemoGender$County <- sub(", Missouri!!Percent", "", ACSDemoGender$County) # prep for merge

# Select percent
ACSDemoRace <- ACSDemo %>%
  dplyr::select(County = V1, BlackAfricanAmericanPercent = V41) %>%
  filter(!str_detect(County, "!!Margin of Error")) %>%
  filter(!str_detect(County, "!!Percent Margin of Error")) %>%
  filter(!str_detect(County, "!!Estimate"))

ACSDemoRace$BlackAfricanAmericanPercent <- sub("%", "", ACSDemoRace$BlackAfricanAmericanPercent) 
# remove "%" in column 

ACSDemoRace <- ACSDemoRace[-c(1), ] # remove first row with variable description

ACSDemoRace$County <- sub(" County, Missouri!!Percent", "", ACSDemoRace$County) # prep for merge
ACSDemoRace$County <- sub(", Missouri!!Percent", "", ACSDemoRace$County) # prep for merge

```

# 6. Merge ACS Data

```{r mergeACS, warning=FALSE, message=FALSE}

# Merge ACS data

ACS_list <- list(ACSSocial, ACSDemoAge, ACSDemoGender, ACSDemoRace, ACSEcon)

ACS <- ACS_list %>%
  reduce(full_join)

```  

# 7. Clean RUCC Data
```{r datacleanRUCC, warning=FALSE, message=FALSE}

# Clean RUCC data

RUCC <- RUCC %>%
  dplyr::select(-FIPS) %>%
  rename(County = County_Name) %>%
  rename(RUCC = RUCC_2013) %>%
  filter(State == "MO") 

RUCC$County <- sub(" County", "", RUCC$County) # remove county in column for merge

```

# 8. Merge Data

```{r mergedata, warning=FALSE, message=FALSE}

Data_list <- list(MICA, ACS, RUCC) 

Data_with_NA <- Data_list %>%
  reduce(full_join, by = "County") %>%
  dplyr::select(-State)

# Add new variable to reescale predictor 
Data <- Data_with_NA %>% 
  na.omit() %>% # remove NAs
  mutate(HouseholdInternetPercent25 = as.numeric(HouseholdInternetPercent)/25) %>%
    mutate(RUCC_cat= case_when(
    RUCC %in% c(1, 2, 3) ~0,
    RUCC %in% c(4, 5, 6) ~1,
    RUCC %in% c(7, 8, 9) ~2)) 

Data$HouseholdInternetPercent <- as.numeric(Data$HouseholdInternetPercent)
Data$HealthInsurancePercent <- as.numeric(Data$HealthInsurancePercent)
Data$HighSchoolPlusPercent <- as.numeric(Data$HighSchoolPlusPercent)
Data$MedianAge <- as.numeric(Data$MedianAge)
Data$FemalePercent <- as.numeric(Data$FemalePercent)
Data$BlackAfricanAmericanPercent <- as.numeric(Data$BlackAfricanAmericanPercent)
Data$HouseholdMedianIncome <- as.numeric(Data$HouseholdMedianIncome)
Data$RUCC_cat <- as.factor(Data$RUCC_cat)

```

# 9. Visualize Data

```{r dataviz, warning=FALSE, message=FALSE}

# Visualize with plots
ggplot(Data, aes(TotalCount))+
    geom_histogram(breaks=c(0,200,400,600,800,1000,1200,1400,1600,1800), color="black", fill="red")+
  labs(title = "Total Counts of Late-Stage Female Breast Cancer by Missouri County (2014-2018)", y = "Frequency", x = "Total Counts of Late-Stage Female Breast Cancer")

ggplot(Data, aes(TotalCount))+
    geom_density()+
    labs(title = "Total Counts of Late-Stage Female Breast Cancer by Missouri County (2014-2018)", y = "Density", x = "Total Counts of Late-Stage Female Breast Cancer")

```

# 10. Table 1

```{r table1, warning=FALSE, message=FALSE}

# Label variables 

label(Data$HealthInsurancePercent) <- "Health Insurance Coverage (percent)"
label(Data$HouseholdInternetPercent) <- "Household Internet Access (percent)"
label(Data$FemalePercent) <- "Female (percent)"
label(Data$MedianAge) <- "Median Age (years)"
label(Data$BlackAfricanAmericanPercent) <- "Black/African American (percent)"
label(Data$HighSchoolPlusPercent) <- "Educational Attainment: High School and Above (percent)"
label(Data$HouseholdMedianIncome) <- "Household Median Income ($)"

Data$RUCC_cat <-
  factor(Data$RUCC_cat, levels = c(0, 1, 2), 
         labels = c("Metropolitan", "Urban", "Rural"))

label(Data$RUCC_cat) <- "Geographic Residence"

table1(~ RUCC_cat + HouseholdInternetPercent + HealthInsurancePercent + FemalePercent + BlackAfricanAmericanPercent + MedianAge + HighSchoolPlusPercent + HouseholdMedianIncome, data=Data, overall="Total")

```

# 11. Test linearity assumption

```{r linearity, warning=FALSE, message=FALSE}

# Box tidwell test
# Test the assumption of linearity for household internet access

Data$HouseholdInternet_times_log_HouseholdInternet <- Data$HouseholdInternetPercent*log(Data$HouseholdInternetPercent)

model2 <- glm(TotalRate ~ HouseholdInternetPercent + Data$HouseholdInternet_times_log_HouseholdInternet, data=Data, family="poisson") 
summary(model2)

# Since the p-value for HouseholdInternet_times_log_HouseholdInternet >.05, we fail to reject the hypothesis that the linearity assumption is not met using the Box Tidwell test and household internet access can be modeled as a linear term.

```

# 12. Test for overdispersion

```{r overdispersion, warning=FALSE, message=FALSE}

# Run Poisson regression for count data
model.PInternet <- glm(TotalRate ~ HouseholdInternetPercent25 + BlackAfricanAmericanPercent + HouseholdMedianIncome + RUCC_cat, family=poisson, data=Data)
summary(model.PInternet) 

# Test for overdispersion by running a negative binomial model
model.1Internet <-  glm.nb(TotalRate ~ HouseholdInternetPercent25 + BlackAfricanAmericanPercent + HouseholdMedianIncome + RUCC_cat, control=glm.control(maxit=50), data = Data)
summary(model.1Internet)

lrtest(model.PInternet, model.1Internet)

# The LR test indicates that the negative binomial model DOES NOT significantly improve the fit to the data and that we have not violated the mean = variance assumption. Therefore the Poisson model adequately fits these data.

```

# 13. Analyze Data

```{r analyzedata, warning=FALSE, message=FALSE}

exp(stats::coefficients(model.PInternet))
exp(stats::confint(model.PInternet))

# Use RSE
dfInternet <-tidy(coeftest(model.PInternet, vcov. = vcovHC(model.PInternet, type = 'HC3'))) 
dfInternet

# Get IRRs and 95% CI based on RSEs
dfInternet$IRR <- exp(dfInternet$estimate)
dfInternet$lowCI <- exp(dfInternet$estimate - 1.96*dfInternet$std.error)
dfInternet$highCI <- exp(dfInternet$estimate + 1.96*dfInternet$std.error)

dfInternet

write_xlsx(dfInternet, "Poisson Results Internet.xlsx")

```

# 14. Test for Effect Modification

```{r effectmod, warning=FALSE, message=FALSE}

model.0 <- glm(TotalRate ~ HouseholdInternetPercent + BlackAfricanAmericanPercent + HouseholdMedianIncome + RUCC_cat, family = poisson, data = Data)

# To test for effect modification, include cross-product interaction term in model (adjust) and compare to model without
model.0ef <- glm(TotalRate ~ HouseholdInternetPercent + BlackAfricanAmericanPercent + HouseholdMedianIncome + RUCC_cat + HealthInsurancePercent*HouseholdInternetPercent, family = poisson, data = Data)
summary(model.0)

# The interaction term is not significant. 
lrtest(model.0, model.0ef)

```
